
@chapter C2Forth - Compiling C for a Forth System

@*
With the development of MPE's VFX Forth family of compilers
that produce optimised native code from high level Forth,
an experiment was performed to use high level Forth as the
assembly language of a C compiler. The C2Forth system is the
result.
@vskip 0pt
@*
The C2Forth system allows you to compile C source to be run
VFX Forth and XC6/7 cross compilers. It consists of a tool
chain that produces Forth source from C source, and a build
harness for the Forth System.
@vskip 0pt

@section Overview


@vskip -6pt
@*
Traditional C compiler chains usually have four stages.
@vskip 0pt


@table @code
@item PreProcess
This step takes a C source file and parses it
 and any included header files to produce a
 single reference source. Macros are expanded at
 this stage and conditional compilation structures
 are evaluated to remove irrelevant information.

@end table



@table @code
@item Compile
This step takes to output from the preprocess
 stage and produces an assembler listing for the
 target CPU.

@end table



@table @code
@item Assemble
The assembler source from the compile stage is
 assembled into a binary form along with any
 relocation information, symbol tables and
 missing dependancy lists.

@end table



@table @code
@item Link
All the binaries for a project (as output from
 the Assemble stage) are merged along with any
 pre-compiled libraries. Dependant symbols from
 an object are resolved with exported symbols
 from the surrounding binaries and libraries to
 form a standalone image.

@end table

@*
MPE has developed a C tool chain which targets a virtual machine
rather than a specific processor. The compiler is based on the LCC
system with the preprocess/assembler/linker derived from other
sources.
@vskip 0pt
@*
The MPE virtual machine design is a dual stack, no register
architecture design which is similar to a conventional Forth
virtual machine. The virtual machine runtime is written in
Forth.
@vskip 0pt

@subsection C Preprocessor


@vskip -4pt
@*
The preprocessor used comes as part of the LCC toolchain.
Only detail changes have been made to this code, mostly to
aid compilation.
@vskip 0pt

@subsection C Compiler


@vskip -4pt
@*
The compiler is a heavily modified derivative of LCC. The LCC
system is designed such that all the processing of C source into
DAG lists as well as tree reduction and analysis are common. Only
the part of the code which translates the final DAG trees into
assembler need retargeting. The "back-end" which performs this job
is usually built with lBurg.
@vskip 0pt
@*
lBurg is a code-generator generator system developed by the same
authors as LCC. Unfortunately the LCC and lBurg combination was
never designed to cope with the kind of architecture our dual
stack VM used. For that reason it became necessary to replace the
"normal" back-end interface in LCC with our own.
@vskip 0pt

@subsection Assembler and Linker


@vskip -4pt
@*
A VFX Forth Kernel replaces the more traditional Assembler and
Linker. The output from the compiler for each file in the project
is passed into the Forth Kernel which has a support harness. The
binary is placed directly into the Forth dictionary and label names
are linked either against each other or against the current Forth
search-order. Therefore libraries and such are not linkable binaries
but Forth definitions built into the system before the converted
code.
@vskip 0pt

@section Current Status and Goals


@vskip -6pt
@*
This tool chain has reached the stage of proving that a good Forth
compiler can be used as a back-end for C. As the usage of C2Forth
increases, the amount of carnal knowledge about the target Forth
system will decrease. Ultimately the goal is to have output source
which is 100% ANS Forth compliant. This is theoretically possible to
achieve - just difficult.
@vskip 0pt
@*
No attempt has been made to either implement a harness or port the
existing one onto anything other than an MPE VFX Forth system.
@vskip 0pt

@section Contents of the distribution


@vskip -6pt

@cartouche
@example
\BIN            C compiler binaries
    CPP         C Preprocessor
    FCC         C Compiler
    OMAKE       Make utility
    C2FPOST     Parsing utility to take FCC output and pad
                it into a form which can be safely TFTP'd
                in 512byte packets.
\DOC            C2F Distribution documentation.
                See C2Forth.pdf
\DOC\LIBC       Documentation on sample C library implementation.
\HARNESS        C2F Forth Target Harness, reference source.
                See C2Forth.pdf for details.
\INCLUDE        C Header files for sample C library implementation.
\LIB            Forth source for C library functions.
\SRC            C source code for tools.
\TESTS          Some simple test programs and a suitable makefile @end example
@end cartouche
